Phylodynamic inference of contact network parameters through approximate Bayesian computation

### 2016-05-30 doc/thesis/intro.tex
A bit more writing

{\color{blue}\uline{
Recall that the primary objective of our work is to develop a statistical
inference method for estimating contact network parameters from transmission
trees. In particular, for a network model with parameters $\theta$ and an input
transmission tree $T$, we are interested in the posterior distribution}
\[
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)}.
\]
\uline{As outlined in} \cref{sec:obj}, \uline{both the likelihood $\Pr(T \mid
\theta)$ and the normalizing constant $\Pr(T)$ are likely computationally
intractable. Hence, rather than computing the posterior distribution
analytically, we will attempt to approximate it using a \defn{Monte Carlo}
approach.
TODO
}}

@@

\subsection{Sequential Monte Carlo}

>>>>

{\color{blue}\uline{
Recall that the primary objective of our work is to develop a statistical
inference method for estimating contact network parameters from transmission
trees. In particular, for a network model with parameters $\theta$ and an input
transmission tree $T$, we are interested in the posterior distribution}
\[
  \Pr(\theta \mid T) = \frac{\Pr(T \mid \theta) \Pr(\theta)}{\Pr(T)}.
\]
\uline{As outlined in} \cref{sec:obj}, \uline{both the likelihood $\Pr(T \mid
\theta)$ and the normalizing constant $\Pr(T)$ are likely computationally
intractable. Hence, rather than computing the posterior distribution
analytically, we will approximate it using a \defn{Monte Carlo} approach.
The fundamental idea behind Monte Carlo methods is succinctly expressed by
\textcite{liu2001theoretical}:
\begin{quote}
    Monte Carlo's view of the world is that any probability distribution $\pi$, 
    regardless of its complexity, can always be \emph{represented} by a
    discrete sample from it. By ``represented'', we mean that any computation
    of expectations using $\pi$ can be replaced to an acceptable degree of
    accuracy by using the empirical distribution resulting from the discrete
    sample.
\end{quote}
In other words, if we are able to sample enough points from a distribution of
interest, we will be able to make reasonably accurate statements about the
distribution itself. For example, the expected value of the distribution can be
estimated by the sample's population mean. Several algorithms exist for
obtaining such a sample, some of which will be discussed further in
\cref{subsec:mfit}. The particular method we apply in this work is \gls{SMC},
which is also known as the \emph{particle filter}. The reason for this name is
that the desired number of samples from the target distribution is decided at
the outset, and that number of ``particles'' \ldots TODO

This is in contrast 
}

\uline{In the
context of this work, the distribution of interest is the posterior
distribution on the parameters of a contact network model, given observed
phylogenetic data in the form of a transmission tree.}

\uline{A typical Bayesian Monte Carlo approach to this problem would be to
calculate the product of the likelihood and the prior for each of the parameter
particles.

}}

@@

\subsection{Sequential Monte Carlo}
\label{subsec:smc}
